---
title: "Tracking COVID-19 Cases Across Different Settlement Types in New York State"
author: "SMiltonDZouGKhalsa"
date: "2022-11-22"
output:
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 2
  html_document:
    toc: yes
    toc_depth: '2'
    df_print: paged
geometry: margin=2.54cm
subtitle: https://github.com/sgm22/MiltonZouKhalsa__ENV872_EDA_FinalProject
fontsize: 12pt
mainfont: Times New Roman
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage
\listoffigures 
\newpage

## Rationale and Research Questions

### Rationale

> Coronavirus disease (COVID-19) is an infectious disease caused by the severe acute 
respiratory syndrome coronavirus 2 (SARS-CoV-2) (Coronavirus, n.d.). The virus is
primarily spread from person to person through respiratory droplets and as such, 
population density plays a role in the transmission of the virus with many experts
suggesting that physical distancing is an effective way to combat its 
spread (CDCMMWR, 2020; What Is Coronavirus?, 2022). Since COVID-19 was declared
a global pandemic in March 2020 by the World Health Organization (WHO) in March 
2020, there has been a resurgence in COVID-19 infections across the globe as a
result of new strains of the virus (CDC, 2020). Policy decisions to protect human
health leverages data that such as case numbers, hospitalizations, and death
resulting from COVID-19 to  project COVID-19 spread (Truelove et al., n.d.). 
Very few epidemiological COVID-19 models incorporate population density explicitly 
even though population density is thought to be an important
factor(Hamidi et al., 2020; Wong & Li, 2020). 

> In New York State, since the beginning of the pandemic there has been over
6 million cases of COVID-19. In January 2022, the state recorded the highest 
number of cases while in April 2020, it recorded the highest record of death 
(Times, 2020). The density of people in New York State varies drastically across 
the state with 43% of the state's population living in the five boroughs of New 
York City (New York State Tracking Program | Tracking | NCEH | CDC, 2019). Of the
62 counties in the State, Kings County is the most populated with over 2 million
residents, while Hamilton County is the least populated with about 4,485 residents 
(New York Population 2022 (Demographics, Maps, Graphs), n.d.). With physical 
distancing playing an important role in reducing the likelihood of infection, 
this paper explores how well absolute population and population density predict
COVID-19 case numbers in the diversely sized New York State. Further, we explore
whether the relationship between absolute population and population density changes 
among different settlement types.

> Settlement types are defined as population centers of humans who have developed 
a long-term community in a specific area (Settlements Overview & Types | What Are Settlements?, n.d.). 
In this study, settlement types are classified as rural, micropolitan, and urban. 
We define rural settlements as counties with a population size of less than
25,000 people, micropolitans as counties with a population size of between 25,000 
and 49,999 people, and an urban settlement as a county with more than 50,000 people
(adapted from USDA ERS - What Is Rural?, n.d.). We hope that by better understanding 
the importance of population size and density on COVID-19 prevalence and how that 
varies across settlement type, we are able to better understand if population density and/or
absolute population can be used as a proxy when formulating policy decisions around COVID-19.


### Research Questions

1. What is the correlation between population density and absolute population 
and the number of COVID-19 cases across New York State?

2. What is the relationship between the predictor variable: settlement types and
the response variable: cumulative COVID-19 cases?

3. What is the correlation between absolute population and population density 
and the number of COVID-19 tests does done? Does the stronger correlated 
relationship vary among settlement types? 

\newpage


**Set-up Working Directory, Import and Merge Data sets**

> In order to answer our research question, we imported data that contains information 
about New York (NY) state population size and density by county in 2022 from 
the U.S. Census Bureau. This information was then combined with data regarding 
number of positive COVID-19 cases, quantity of COVID-19 tests conducted and other
relevant information collected by the New York Department of Health from March 
1, 2020 to November 22, 2022. 

```{r, message=FALSE, warning=FALSE}

################# Set working directory ################################

getwd() #Check working directory

#Load packages

library(rvest)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library (lubridate)
library (dplyr)
library(sf)
library (leaflet)
library(mapview)
library(AER)
library (corrplot)
library(RColorBrewer)
library(knitr)
library(MASS)
library(kableExtra)

#Set theme

my_theme <- theme_bw(base_size = 12) + 
  theme(axis.text = element_text(color = "black"), 
      legend.position = "top", legend.justification = "center") +
  theme(plot.title = element_text(hjust = 0.5))
theme_set(my_theme)


#Import NY State county level COVID data (March 1, 2020 to November 22, 2022)
#from NY Health.

ny_covid_dat <- read.csv("./Data/Raw_Data/NY_COVID_County_Level.csv")

ny_covid_dat #View Data

#Import NY population data (2022)

ny_pop_dat <- read.csv("./Data/Raw_Data/NY_County_Level_Population_Data.csv"
                       ,stringsAsFactors = TRUE)

ny_pop_dat

#Merge data sets

combined_dat_ny_covid_pop <- ny_pop_dat %>% 
  left_join(ny_covid_dat, by = "County")

## Add column for to classify counties into settlement types

combined_dat_ny_covid_pop <- combined_dat_ny_covid_pop %>%
  mutate(settlement_type = case_when(pop2022<=25000 ~ 'Rural',
                                     25000<=pop2022 & pop2022<50000 ~ 'Micropolitan',
                                     pop2022>= 50000 ~ 'Urban'))

summary(combined_dat_ny_covid_pop) # Check data for new columns

```

\newpage


## Dataset Information 

> We acquired our 2020 to 2022 COVID-19 data from the New York Department of Health 
and population data from the U.S. Census Bureau on November 24th, 2022. County 
names did not match between the census data and NY Department of Health data. 
We wrangled our data by removing the word "county" from each county in the census 
data. In addition to the 62 counties, the NY Health Department broke counties down 
into smaller sections resulting in more than 62 counties. Since we only want to
work with the original counties, we left-merged the NY health data onto the county 
data on November 24th, 2022. We are using the merged version as our main data source.
Additionally on November 24th, 2022, we uploaded the geospatial shape file that was
incorporated in class for ENV872.Using the shape file, we selected for
counties in New York State. Data types and variable types for the combined data 
were set on November 27th, 2022.

Data Information:
https://www.census.gov/library/stories/state-by-state/new-york-population-change-between-census-decade.html
COVID-19 Variant Data | Department of Health (ny.gov) – (March 1, 2020,  to November 22, 2022)
cb_2018_us_county_20m.shp state FIPS code 36

**Keywords**

ENV872, COVID-19, COVID, virus, virus spread, New York, New York State

```{r, echo=FALSE}

meta_data_tab <- read.csv("./Data/Meta_Data/meta_data_table.csv")

print(meta_data_tab)
```

\newpage

## Exploratory Analysis and Data Visualization {.tabset}

> The merged data set was exploring for missing values and abnormal values using the R functions, "head", "summary" and "dim". Further, to better understand the distribution of the variables that we use in our analysis, we plotted a series of histograms. The graphical exploration of the distribution of the variables showed that the variables, "sum_cumulative_no_of_positive", "sum_cumulative_no_of_tests", "pop2022" and "density_permilessquared" are right-skewed. That is, most of the data points were located on the left side of the histograms plotted for each variable. For improved visualization, we log transformed the variables to get their distributions closer to a normal distribution. After each log-transformation, we carried out Shapiro–Wilk tests to check the normality of the transformed distributions. The p-value of each log-transformed distribution was less than 0.05 indicating that the evidence supported the fact that log-transformed distributions were still not normally distributed. We also explored and visualized the distribution of the absolute population and population density across the various settlement types using box plots.

### Tabular Exploration

```{r, message=FALSE, warning=FALSE}

head(combined_dat_ny_covid_pop)
summary(combined_dat_ny_covid_pop) #View Data
dim(combined_dat_ny_covid_pop)


#View mean COVID cases and mean COVID test conducted by settlement types


collapsed_dat_settlement_types <- combined_dat_ny_covid_pop %>%
  group_by(settlement_type) %>%
  summarise(Mean_COVID_cases = mean(sum_cumulative_no_of_positive),
            Mean_Tests_Conducted = mean(sum_cumulative_no_of_tests),
            Mean_Absolute_Population = mean(pop2022),
            Mean_Settlement_Density = mean (density_permilessquared))

print(collapsed_dat_settlement_types) #View table

```

### Graphical Exploration

```{r, message=FALSE, warning=FALSE, fig.cap= "Cumulative Number of Positive COVID-19 Tests", fig.align='left', fig.height= 2.5 }

# Visualizing Distribution of Variables

###################### COVID-19 Cases ##############################

ggplot(data = combined_dat_ny_covid_pop, aes(sum_cumulative_no_of_positive))+
  geom_histogram() +
  xlab ("Cumulative Number of Positive COVID-19 Tests") +
  ggtitle("Histogram of Cumulative Number of Positive COVID-19 Tests")

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Log of Cumulative Number of Positive COVID-19 Tests", fig.align='left', fig.height= 2.5 }

## Log Transform

ggplot(data = combined_dat_ny_covid_pop,
       aes(log(sum_cumulative_no_of_positive))) +
  geom_histogram() +
  xlab ("Log of Cumulative Number of Positive COVID-19 Tests") +
  ggtitle("Histogram of Log of Cumulative Number of Positive COVID-19 Tests")

shapiro.test(log(combined_dat_ny_covid_pop$sum_cumulative_no_of_positive)) 
#Check normality

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Cumulative Number of COVID-19 Tests Conducted", fig.align='left', fig.height= 2.5 }

####################### Number of COVID Tests Conducted #######################

ggplot(data = combined_dat_ny_covid_pop, aes(sum_cumulative_no_of_tests))+
  geom_histogram() +
  xlab ("Cumulative Number of COVID-19 Tests Conducted") +
  ggtitle("Histogram of Cumulative Number of COVID-19 Tests Conducted")

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Log of Cumulative Number of Positive COVID-19 Tests Conducted", fig.align='left', fig.height= 2.5 }

## Log Transform

ggplot(data = combined_dat_ny_covid_pop, aes(log(sum_cumulative_no_of_tests))) +
  geom_histogram() +
  xlab ("Log of Cumulative Number of COVID-19 Tests Conducted") +
  ggtitle("Histogram of Log of Cumulative Number of COVID-19 Tests Conducted")

shapiro.test(log(combined_dat_ny_covid_pop$sum_cumulative_no_of_tests))
#Check normality

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Population of Counties Across New York State", fig.align='left', fig.height= 2.5 }

######################### County Population ##########################

ggplot(data = combined_dat_ny_covid_pop, aes(pop2022)) +
  geom_histogram() +
  xlab("Absolute Population Size of Counties Across New York State 2022") +
  ggtitle ("Histogram of Population of Counties Across New York State")

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Log of Population of Counties Across New York State", fig.align='left', fig.height= 2.5 }

## Log Transform

ggplot(data = combined_dat_ny_covid_pop, aes(log(pop2022))) +
  geom_histogram() +
  xlab("Log of Absolute Population of Counties Across New York State 2022 ") +
  ggtitle ("Log of Histogram of Population of Counties Across New York State")

shapiro.test(log(combined_dat_ny_covid_pop$pop2022)) #Check normaliity

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Density of Counties Across New York State", fig.align='left', fig.height= 2.5 }

################################## County Density ############################

ggplot(data = combined_dat_ny_covid_pop, aes(density_permilessquared)) +
  geom_histogram() +
  xlab("Density of Counties Across New York State") +
  ggtitle (" Histogram of Density of Counties Across New York State")

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Log of Density of Counties Across New York State", fig.align='left', fig.height= 2.5 }

## Log Transform

ggplot(data = combined_dat_ny_covid_pop, aes(log(density_permilessquared))) +
  geom_histogram() +
  xlab("Log of Density of Counties Across New York State") +
  ggtitle (" Log of Histogram of Density of Counties Across New York State")

shapiro.test(log(combined_dat_ny_covid_pop$density_permilessquared))
#Check normality

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Log of Absolute Population Size of Counties by Settlement Type", fig.align='left', fig.height= 2.5 }
########### Visualize Distribution of Predictor Variables by Settlement Type

ggplot(data = combined_dat_ny_covid_pop, aes(x= settlement_type,
                                             y= log(density_permilessquared),
                                             fill = settlement_type))+
  geom_boxplot() +
  ggtitle ("Log of Absolute Population Size of Counties by Settlement Type") +
  xlab("Settlement Types") +
  ylab("Log of Absolute Population Size 2022")

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Log of Population Density of Counties by Settlement Type", fig.align='left', fig.height= 2.5 }

ggplot(data = combined_dat_ny_covid_pop,
       aes(x= settlement_type, y= log(pop2022), fill = settlement_type))+
  geom_boxplot() +
    ggtitle ("Log of Population Density of Counties by Settlement Type") +
  xlab("Settlement Types") +
  ylab("Log of Absolute Population Size 2022")

```


### Spatial Context of Data

```{r, message=FALSE, warning=FALSE}

# Map of New York County Population Levels
 
# The EPSG code of the New York counties dataset is 4269.
#This is a geographic coordinate reference system, so it uses
#angular coordinate units. This CRS (4269) is associated with the datum NAD83.

uscounty <- st_read('./Data/Spatial_Files/cb_2018_us_county_20m.shp',
                    stringsAsFactors = TRUE) %>%
 filter(STATEFP == 36)
 
# Reveal the names of the columns
colnames(combined_dat_ny_covid_pop)
 
# Join the flow data to our NWIS gage location spatial dataframe.
combined_dat_join <- merge(x = uscounty,
                          y = combined_dat_ny_covid_pop,
                          by.x = "NAME",
                          by.y = "County")
 
# Show the column names of the joined dataset.
colnames(combined_dat_join)
 
# Show the dimensions of this joined dataset.
dim(combined_dat_join)

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Population of New York Counties", fig.align='center', fig.height= 4, fig.width= 5}

# Map the population of each New York county in 2022.
ggplot() +
  geom_sf(data = combined_dat_join, aes(fill = log(pop2022))) +
  scale_fill_gradientn(colours = terrain.colors(10)) +
  ggtitle("COVID-19 in New York",
          subtitle = "Population by County in 2022") +
  labs(fill = 'Log of Variable')

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Settlement Types of New York Counties", fig.align='center', fig.height= 4, fig.width= 5}
 
# Map the settlement types of each New York county in 2022.
ggplot() +
  geom_sf(data = combined_dat_join, aes(fill = settlement_type)) +
  ggtitle("COVID-19 in New York",
          subtitle = "Settlement Types by County in 2022") +
  labs(fill = 'Settlement Types')

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Cumulative Number of Tests Administered in Each New York County", fig.align='center', fig.height= 4, fig.width= 5}
 
# Map the sum of the cumulative number of tests administered in each
#New York county in 2022.

ggplot() +
  geom_sf(data = combined_dat_join,
          aes(fill = log(sum_cumulative_no_of_tests))) +
  scale_fill_gradientn(colours = terrain.colors(10)) +
  ggtitle("COVID-19 in New York",
          subtitle = "Cumulative Number of Tests Administered") +
  labs(fill = 'Log of Variable')

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Cumulative Number of Positive Test Results in Each New York County", fig.align='center', fig.height= 4, fig.width= 5}
 
# Map the sum of the cumulative number of positive test results in each
#New York county in 2022.

ggplot() +
  geom_sf(data = combined_dat_join,
          aes(fill = log(sum_cumulative_no_of_positive))) +
  scale_fill_gradientn(colours = terrain.colors(10)) +
  ggtitle("COVID-19 in New York",
          subtitle = "Cumulative Number of Positive Test Results in 2022") +
  labs(fill = 'Log of Variable')

```

**Export Processed data**

```{r, message=FALSE, warning=FALSE}

getwd() #Check working directory

#Full processed data set

write.csv2(combined_dat_ny_covid_pop,
           "./Data/Processed_Data/Combined_NY_COVID_Pop.csv")

#Collapsed data set

write.csv2(collapsed_dat_settlement_types,
           "./Data/Processed_Data/collapsed_dat_settlement_types.csv")
```


\newpage


## Analysis {.tabset}

> As a central question of this study is "How  well does population density and
absolute population predict COVID-19 cases?", the predictor variables are 
population density per miles squared and absolute population while the outcome 
variable is sum cumulative positive number of positive COVID-19 tests which is 
being used as a proxy for COVID-19 cases. Our unit of analysis is counties and 
settlement types depending on the primary objective of each sub-question. 
Settlement types are classified based on our definitions of rural, micropolitan,
and urban settlement types. Our hypotheses for each study question
is indicated below.

**Question 1:  What is the correlation between population density and absolute population and the number of COVID-19 cases (cumulative number of positive tests) across New York State?**

> Prior to the correlation analysis, the normality of absolute population, population 
density, and COVID-19 cases were examined and were determined to not be normally 
distributed (see exploratory analysis). The variables were log-transformed but
were showed to still not be normally distributed after a Shapiro-Wilk test of
normality was performed. As such, a Spearman correlation analysis for non-parametric 
data was carried out. The classification of the strength of the relationship was 
determined based on the value of r which ranges from 0 to 1 where r = 0 indicates
no association and r = -1 or +1.

#### Absolute Population and COVID-19 Cases

*Hypotheses*

Null Hypothesis: As the ranks of absolute population change, the ranks of the number of COVID-19 cases do not change. 

Alternative Hypothesis: As the ranks of absolute population change, the ranks of the number of COVID-19 cases change.


*Assumptions:*

1. The relationship between both variables is monotonic.


```{r, message=FALSE, warning=FALSE}

# Spearman rank correlation test as distributions are skewed


correlation_results_absol_pop <-
  cor.test(x=log(combined_dat_ny_covid_pop$pop2022),
           y=log(combined_dat_ny_covid_pop$sum_cumulative_no_of_positive),
           method = "spearman")
correlation_results_absol_pop

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Correlation Between Absolute Population Across Counties and COVID-19 Cases", fig.align='default', fig.height= 2.5 }

# Visualize plot (Log-transformed for better visualization)

ggplot(combined_dat_ny_covid_pop, aes(x = log(pop2022),
                                      y = log(sum_cumulative_no_of_positive)))+
  geom_point() +
  ylab ("Log of COVID-19 Cases") +
  xlab ("Log of Absolute Population Across Counties") +
  ggtitle ("The Correlation Between Absolute Population 
           Across Counties and COVID-19 Cases") +
  stat_smooth (method = 'lm') +
  annotate("text",x=9,y=20, label=(paste0("R=0.99")))
  
```

### Population Density and COVID-19 Cases

*Hypotheses:*

Null Hypothesis: As the ranks of population density change, the ranks of the number of COVID-19 cases do not change.

Alternative Hypothesis: As the ranks of population density change, the ranks of the number of COVID-19 cases change.

*Assumptions:*

1. The relationship between both variables is monotonic.

```{r, message=FALSE, warning=FALSE}

# Spearman rank correlation test as distributions are skewed


correlation_results_pop_density <- cor.test(x=log(combined_dat_ny_covid_pop$density_permilessquared),
         y=log(combined_dat_ny_covid_pop$sum_cumulative_no_of_positive),
                                          method = "spearman")
correlation_results_pop_density

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Correlation Between Population Densities Across Counties and COVID-19 Cases", fig.align='left', fig.height= 2.5 }

# Visualize plot (Log-transformed for better visualization)

ggplot(combined_dat_ny_covid_pop, aes(x = log(density_permilessquared),
                                      y = log(sum_cumulative_no_of_positive)))+
  geom_point() +
  xlab (" Log of Population Density (per miles squared)") +
  ylab ("Log of COVID-19 Cases") +
  ggtitle ("The Correlation Between Population Densities
           Across Counties and COVID-19 Cases") +
  geom_smooth (method = 'lm')+
  annotate("text",x=3,y=20, label=(paste0("R=0.89")))

```


**Question 2:  What is the relationship between the predictor variable: settlement types and the response variable: cumulative COVID-19 cases?**

#### Negative Binomial Regression

> In addition to the correlation analysis, we carried out a negative binomial regression to further understand the relationship between settlements types (determined by levels of absolute population) and cumulative COVID-19 cases. We used a negative binomial regression as based on the exploratory plots, we believed that the outcome variable can be classified as over-dispersed count data. We check our assumption that the outcome variable,"sum_cumulative_no_of_positive", was overdispersed by carrying out a Poisson regression and found that the negative binomial regression was a better fit for the data. Further, we carried out a Chi-Square Goodness of Fit test for our negative binomial regression model with residual deviance of 75.355 on 59 degrees of freedom and the p-value obtained was 0.07. This means that the evidence is more supportive of the fact that the model is a good fit.

*Hypotheses:*

Null Hypothesis: There is no trend between the predictor variable (settlement types) and outcome variable (cumulative COVID-19 cases).

Alternative Hypothesis: There is a trend between the response (settlement types) and outcome variables (cumulative COVID-19 cases).

*Assumptions:*

1. The response variable (COVID cases) is a count variable.
2. Each observation of data is independent of each other.
3. The mean and variance are not equal.


```{r, message=FALSE, warning=FALSE,}

## Negative Binomial Regression

summary (modelnb <-
           glm.nb(sum_cumulative_no_of_positive ~ factor(settlement_type),
                  data = combined_dat_ny_covid_pop))


par (mfrow = c(2,2))
plot(modelnb) #Check fit

# Test of Fitness

pchisq(q=modelnb$deviance, df=modelnb$df.residual, 
lower.tail=FALSE)  #Check model fit


## Check our assumption that a negative binomial is a better model than the
#poisson model

summary (modelpoisson <-
           glm(sum_cumulative_no_of_positive ~ factor(settlement_type),
               family = "poisson", data = combined_dat_ny_covid_pop))

## Summary inference: Residual deviance is much larger than degrees of
#freedom (DF). Negative binomial plot better.

```


**Question 3: What is the correlation between absolute population and population density and the number of COVID-19 tests done?**

> Similar to question 1, prior to the correlation analysis, the normality of absolute 
population, population density, and COVID-19 cases were examined and were determined to not be normally 
distributed (see exploratory analysis). The variables were log-transformed but
were showed to still not be normally distributed after a Shapiro-Wilk test of
normality was performed. As such, a Spearman correlation analysis for non-parametric 
data was carried out. The classification of the strength of the relationship was 
determined based on the value of r which ranges from 0 to 1 where r = 0 indicates
no association and r = -1 or +1.

#### Absolute Population and COVID-19 Cases


*Hypotheses:*

Null Hypothesis: As the ranks of absolute population change, the ranks of the number of COVID-19 tests conducted do not change. 

Alternative Hypothesis:  As the ranks of absolute population change, the ranks of the number of COVID-19 tests conducted change.

*Assumptions*

1. The relationship between both variables is monotonic.

```{r, message=FALSE, warning=FALSE}

# Spearman rank correlation test as distributions are skewed


correlation_results_absol_pop <-
  cor.test(
    x = log(combined_dat_ny_covid_pop$pop2022),
    y = log(combined_dat_ny_covid_pop$sum_cumulative_no_of_tests),
    method = "spearman")
correlation_results_absol_pop

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Correlation Between Absolute Population Across Counties and COVID-19 Tests Conducted", fig.align='left', fig.height= 2.5}

# Visualize plot (Log-transformed for better visualization)

ggplot(combined_dat_ny_covid_pop,
       aes(x = log(pop2022),
           y = log(sum_cumulative_no_of_tests)))+
  geom_point() +
  ylab ("Log of COVID-19 Tests Conducted") +
  xlab ("Log of Absolute Population Across Counties") +
  ggtitle ("The Correlation Between Absolute Population 
           Across Counties and COVID-19 Tests Conducted") +
  stat_smooth (method = 'lm') +
  annotate("text",x=9,y=22, label=(paste0("R=0.97")))
  
```

#### Population Density and COVID-19 Cases


*Hypotheses:*

Null Hypothesis: As the ranks of population density change, the ranks of the number of COVID-19 tests conducted do not change.

Alternative Hypothesis: As the ranks of population density change, the ranks of the number of COVID-19 tests conducted change.

*Assumptions:*

1. The relationship between both variables is monotonic.

```{r, message=FALSE, warning=FALSE}

# Spearman rank correlation test as distributions are skewed


correlation_results_pop_density <- cor.test(
  x= log(combined_dat_ny_covid_pop$density_permilessquared),
  y = log(combined_dat_ny_covid_pop$sum_cumulative_no_of_tests),
  method = "spearman")
correlation_results_pop_density

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Correlation Between Absolute Population Densities Across Counties and COVID-19 Tests Conducted", fig.align='left', fig.height= 2.5}

# Visualize plot (Log-transformed for better visualization)

ggplot(combined_dat_ny_covid_pop, aes(x = log(density_permilessquared), y = log(sum_cumulative_no_of_tests)))+
  geom_point() +
  xlab (" Log of Population Density (per miles squared)") +
  ylab ("Log of COVID-19 Tests Conducted") +
  ggtitle ("The Correlation Between Population Densities
           Across Counties and COVID-19 Tests Conducted") +
  geom_smooth (method = 'lm')+
  annotate("text",x=3,y=23, label=(paste0("R=0.86")))

```


**Question 3b: Does the stronger correlated relationship vary among settlement types?**

#### Absolute Population and Settlement Types

```{r, message=FALSE, warning=FALSE}

### Correlation Plots by Settlement Type

##Urban

Urban_dat <- combined_dat_ny_covid_pop %>% 
  filter(settlement_type=="Urban")


correlation_results_absolute_pop_urban <-
  cor.test(x= log(Urban_dat$pop2022),
           y = log(Urban_dat$sum_cumulative_no_of_tests), method = "spearman")

correlation_results_absolute_pop_urban

## Micropolitan

Micropolitan_dat <- combined_dat_ny_covid_pop %>% 
  filter(settlement_type=="Micropolitan")


correlation_results_absolute_pop_microp <-
  cor.test(x= log(Micropolitan_dat$pop2022),
           y = log(Micropolitan_dat$sum_cumulative_no_of_tests),
          method = "spearman")

correlation_results_absolute_pop_microp

##Rural

Rural_dat <- combined_dat_ny_covid_pop %>% 
  filter(settlement_type=="Rural")


correlation_results_absolute_pop_rural <- 
  cor.test(x= log(Rural_dat$pop2022),
           y = log(Rural_dat$sum_cumulative_no_of_tests), method = "spearman")

correlation_results_absolute_pop_rural

```

#### Visualization of 3b

```{r, message=FALSE, warning=FALSE, fig.cap= "Log of Cumulative COVID-19 Cases Across Settlement Types", fig.align='left', fig.height= 2.5}

ggplot(combined_dat_ny_covid_pop, aes(x = log(pop2022),
                                      y = log(sum_cumulative_no_of_positive)))+
  geom_point() + 
  ylab ("Log of Cumulative Number of COVID-19 Cases") +
  xlab ("Absolute Population (2022)")+
  facet_wrap(vars(settlement_type)) 
 
```


\newpage

## Summary and Conclusions {.tabset}

#### General Overview of Data

>The most populous county in New York State according to the information we obtained from the US Census Bureau is Kings County which has a population of over 2.7 million people. Meanwhile, the least populous county was Hamilton county with a little over 5000 residents. Despite Kings County have the most residents in the state, New York City was the most densely populated county - with 75,739 residents per squared miles. In addition to being the least populated county, Hamilton was the least dense of the 62 counties in New York State.
  
#### Correlation between Absolute Population and Population Density with COVID-19 Cases

>We set out to answer our central question, "How well absolute population and population density predict
COVID-19 case numbers in the diversely sized New York State?" by first examining the correlation between population density 
and absolute population and the number of COVID-19 cases across New York. In both cases
of population density and absolute population, after performing a Spearman correlation
for both variables we had a p-value of 2.2e-16, which suggests that the correlated relationship was statistically significant. This means we  have sufficient evidence to reject our null hypothesis that as the ranks of absolute populationand population density change, the ranks of the number of COVID-19 cases do not 
change and instead accept our alternative hypothesis that as the ranks of absolute
population and population density change, the ranks of the number of COVID-19 cases
does change. It is also demonstrated in the visualized plots that there is indeed 
a positive correlation between the variables as the R-squared value for the population 
density plot was 0.89 and 0.99 for the absolute population plot. Both show a 
strong correlation, but it is stronger with absolute population.

#### Association between COVID-19 Cases and Settlement Types

>In the second question, we examined the relationship between our predictor 
variable (settlement types) and response variable (cumulative COVID-19 cases). 
We did not control for potential confounding factors which limits the strength of
the conclusions we are able to draw from this analysis. After performing a negative
binomial regression, we observed that the variable level, rural, has a coefficient 
of -1.1844 which is associated with a p-value of 0.127. This means that when compared
to the reference settlement level, micropolitan, the difference in the logs of the 
expected counts is expected to be 1.1844 lower for rural when compared to micropolitan.
However, with a p-value of 0.127, the evidence is more supportive of the fact that this 
relationship is not significant. Meanwhile, we observed that when compared to the
reference level, micropolitan, the difference in the logs of expected counts is 
expected to be 7.931 higher for urban settlements. The p-value obtained for this
association is less than 0.05 indicating that the evidence is more supportive of 
the fact that this association is statistically significant. 

#### Correlation between Absolute Population and Population Density with Number of COVID-19 Tests Conducted

>In our last question, we examined the correlation between absolute population 
and population density and the number of COVID-19 tests conducted in New York. 
Similar to the results of our first question, after performing a Spearman rank 
correlation test with both absolute population and population density we found
that both had a p-value of 2.2e-16, indicating that the correlated relationship 
was statistically significant.Therefore, we have sufficient evidence to reject 
both null hypotheses that as ranks of absolute populationand population density 
change, the the ranks of the number of COVID-19 tests conducted do not change.
Instead, we accept our alternative hypotheses that as ranks of absolute population
and population density change, the ranks of the number of COVID-19 tests 
conducted does change. This is also shown in our visualized plots, 
with the absolute population plot having an R-squared 
value of 0.97 and the population density plot having an R-squared value of 0.86.
Both show a strong correlation, but it is stronger with absolute population.

>After determining that the correlation between absolute population and number of COVID-19 tests conducted is stronger, we examined within those parameters to see if this correlation varies among settlement types. We performed the Spearman rank correlation test on all three settlement types. Urban settlements had a p-value of 2.2e-16 and an R-squared value of 0.96, micropolitan settlements had a p-value of 0.002 and an R-squared value of 0.77, and rural settlements had a p-value of 1 and an R-squared value of 0.5. Two of the p-values show that the correlated relationships are significant, and rural didn’t have a significant correlated relationship. The correlation is strongest with urban settlements. However, we do acknowledge that the rural settlements only had 3 data points, which is much fewer than the amount of data of micropolitan and urban settlements.

## Conclusion

>In conclusion, we do see a strong correlation between both absolute population
and population density and the number of COVID-19 cases in New York, as well as
the relationship between both absolute population and population density and 
the number of COVID-19 tests conducted in New York. The absolute
population correlation to the number of COVID-19 tests conducted was the stronger relationship. It is important for us to know the correlation between COVID-19 cases and settlement type so we can be better informed about what factors influence the spread of infectious diseases for the future. 

\newpage

## References

1. CDC. (2020, February 11). COVID-19 and Your Health. Centers for Disease Control and Prevention. https://www.cdc.gov/coronavirus/2019-ncov/your-health/reinfection.html

2. CDCMMWR. (2020). Geographic Differences in COVID-19 Cases, Deaths, and Incidence—United States, February 12–April 7, 2020. MMWR. Morbidity and Mortality Weekly Report, 69. https://doi.org/10.15585/mmwr.mm6915e4

3. Coronavirus. (n.d.). Retrieved March 14, 2022, from https://www.who.int/westernpacific/health-topics/coronavirus

4. Hamidi, S., Sabouri, S., & Ewing, R. (2020). Does Density Aggravate the COVID-19 Pandemic? Journal of the American Planning Association, 86(4), 495–509. https://doi.org/10.1080/01944363.2020.1777891

5. New York Population 2022 (Demographics, Maps, Graphs). (n.d.). Retrieved December 10, 2022, from https://worldpopulationreview.com/states/new-york-population

6. New York State Tracking Program | Tracking | NCEH | CDC. (2019, February 20). https://www.cdc.gov/nceh/tracking/profiles/New_York_State_Profile.htm

7. Settlements Overview & Types | What are Settlements? - Video & Lesson Transcript. (n.d.). Study.Com. Retrieved December 10, 2022, from https://study.com/learn/lesson/settlements-overview-types.html

8. Times, T. N. Y. (2020, April 1). New York Coronavirus Map and Case Count. The New York Times. https://www.nytimes.com/interactive/2021/us/new-york-covid-cases.html

9. Truelove, S., Smith, C. P., Qin, M., Mullany, L. C., Borchering, R. K., Lessler, J., Shea, K., Howerton, E., Contamin, L., Levander, J., Kerr, J., Hochheiser, H., Kinsey, M., Tallaksen, K., Wilson, S., Shin, L., Rainwater-Lovett, K., Lemairtre, J. 

10. C., Dent, J., … Viboud, C. (n.d.). Projected resurgence of COVID-19 in the United States in July—December 2021 resulting from the increased transmissibility of the Delta variant and faltering vaccination. ELife, 11, e73584. https://doi.org/10.7554/eLife.73584

11. USDA ERS - What is Rural? (n.d.). Retrieved November 27, 2022, from https://www.ers.usda.gov/topics/rural-economy-population/rural-classifications/what-is-rural.aspx

12. What Is Coronavirus? (2022, July 29). https://www.hopkinsmedicine.org/health/conditions-and-diseases/coronavirus

13. Wong, D. W. S., & Li, Y. (2020). Spreading of COVID-19: Density matters. PLOS ONE, 15(12), e0242398. https://doi.org/10.1371/journal.pone.0242398


