---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
    toc: true
    toc_depth: 2
geometry: margin=2.54cm
title: "Tracking COVID-19 Cases Across Different Settlement Types in New York State"
subtitle: "https://github.com/sgm22/MiltonZouKhalsa__ENV872_EDA_FinalProject"
author: "SMiltonDZouGKhalsa"
date: "2022-11-22"
fontsize: 12pt
mainfont: Times New Roman
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

# Rationale and Research Questions

## Rationale

Coronavirus disease (COVID-19) is an infectious disease caused by the severe acute 
respiratory syndrome coronavirus 2 (SARS-CoV-2) (Coronavirus, n.d.). The virus is
primarily spread from person to person through respiratory droplets and as such, 
population density plays a role in the transmission of the virus with many experts
suggesting that physical distancing is an effective way to combat its 
spread (CDCMMWR, 2020; What Is Coronavirus?, 2022). Since COVID-19 was declared
a global pandemic in March 2020 by the World Health Organization (WHO) in March 
2020, there has been a resurgence in COVID-19 infections across the globe as a
result of new strains of the virus (CDC, 2020). Policy decisions to protect human
health leverages data that such as case numbers, hospitalizations, and death
resulting from COVID-19 to  project COVID-19 spread (Truelove et al., n.d.). 
Very few epidemiological COVID-19 models incorporate population density explicitly 
even though population density is thought to be an important
factor(Hamidi et al., 2020; Wong & Li, 2020). 

In New York State, since the beginning of the pandemic there has been over
6 million cases of COVID-19. In January 2022, the state recorded the highest 
number of cases while in April 2020, it recorded the highest record of death 
(Times, 2020). The density of people in New York State varies drastically across 
the state with 43% of the state's population living in the five boroughs of New 
York City (New York State Tracking Program | Tracking | NCEH | CDC, 2019). Of the
62 counties in the State, Kings County is the most populated with over 2 million
residents, while Hamilton County is the least populated with about 4,485 residents 
(New York Population 2022 (Demographics, Maps, Graphs), n.d.). With physical 
distancing playing an important role in reducing the likelihood of infection, 
this paper explores how well absolute population and population density predict
COVID-19 case numbers in the diversely sized New York State. Further, we explore
whether the relationship between absolute population and population density changes 
among different settlement types.

Settlement types are defined as population centers of humans who have developed 
a long-term community in a specific area (Settlements Overview & Types | What Are Settlements?, n.d.). 
In this study, settlement types are classified as rural, micropolitan, and urban. 
We define rural settlements as counties with a population size of less than
25,000 people, micropolitans as counties with a population size of between 25,000 
and 49,999 people, and an urban settlement as a county with more than 50,000 people
(adapted from USDA ERS - What Is Rural?, n.d.). We hope that by better understanding 
the importance of population size and density on COVID-19 prevalence and how that 
varies across settlement type, we are able to better understand if population density
can be used as a proxy when formulating policy decisions around COVID-19.


## Research Questions

1. What is the correlation between population density and absolute population 
and the number of COVID-19 cases across New York State?

2. What is the relationship between the predictor variable: settlement types and
the response variable: cumulative COVID-19 cases?

3. What is the correlation between absolute population and population density 
and the number of COVID-19 tests does done? Does the stronger correlated 
relationship vary among settlement types? 


\newpage

In order to answer our research question, we imported data that contains information 
about New York (NY) state population size and density by county in 2022 from 
the U.S. Census Bureau. This information was then combined with data regarding 
number of positive COVID-19 cases, quantity of COVID-19 tests conducted and other
relevant information collected by the New York Department of Health from March 
1, 2020 to November 22, 2022. More information on the files used can be found on
the README file linked below.

### Set-up Working Directory, Import and Merge Datasets

```{r, message=FALSE, warning=FALSE}

################# Set working directory ################################

getwd() #Check working directory

#Load packages

library(rvest)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library (lubridate)
library (dplyr)
library(sf)
library (leaflet)
library(mapview)
library(AER)
library (corrplot)
library(RColorBrewer)
library(knitr)
library(MASS)

#Set theme

my_theme <- theme_bw(base_size = 12) + 
  theme(axis.text = element_text(color = "black"), 
      legend.position = "top", legend.justification = "center") +
  theme(plot.title = element_text(hjust = 0.5))
theme_set(my_theme)


#Import NY State county level COVID data (March 1, 2020 to November 22, 2022) from NY Health.

ny_covid_dat <- read.csv("./Data/Raw_Data/NY_COVID_County_Level.csv")

ny_covid_dat #View Data

#Import NY population data (2022)

ny_pop_dat <- read.csv("./Data/Raw_Data/NY_County_Level_Population_Data.csv",stringsAsFactors = TRUE)

ny_pop_dat

#Merge data sets

combined_dat_ny_covid_pop <- ny_pop_dat %>% 
  left_join(ny_covid_dat, by = "County")

## Add column for to classify counties into settlement types

combined_dat_ny_covid_pop <- combined_dat_ny_covid_pop %>%
  mutate(settlement_type = case_when(pop2022<=25000 ~ 'Rural',
                                     25000<=pop2022 & pop2022<50000 ~ 'Micropolitan',
                                     pop2022>= 50000 ~ 'Urban'))

summary(combined_dat_ny_covid_pop) # Check data for new columns

```

\newpage

# Dataset Information

We acquired our 2020 to 2022 COVID-19 data from the New York Department of Health and population data from the U.S. Census Bureau on November 24th, 2022. County names did not match between the census data and NY Department of Health data. We wrangled our data by removing the word "county" from each county in the census data. In addition to the 62 counties, the NY Health Department broke counties down into smaller sections resulting in more than 62 counties. Since we only want to work with the original counties, we left-merged the NY health data onto the county data on November 24th, 2022. We are using the merged version as our main data source. Additionally on November 24th, 2022, we uploaded the state map file that was incorporated in class for ENV872. Lastly, we added spatial files and updated the data types and variable types for the combined data set on November 27th, 2022.

Data Information:
https://www.census.gov/library/stories/state-by-state/new-york-population-change-between-census-decade.html
COVID-19 Variant Data | Department of Health (ny.gov) – (March 1, 2020,  to November 22, 2022)
cb_2018_us_county_20m.shp state FIPS code 36

## Keywords

ENV872, COVID-19, COVID, virus, virus spread, New York, New York State

```{r, echo=FALSE}

meta_data_tab <- read.csv("./Data/Meta_Data/meta_data_table.csv")

kable(meta_data_tab, caption = "COVID-19 Metadata")

```

\newpage

## Exploratory Analysis and Data Visualization

### Tabular Exploration

```{r, message=FALSE, warning=FALSE}

head(combined_dat_ny_covid_pop)
summary(combined_dat_ny_covid_pop) #View Data
dim(combined_dat_ny_covid_pop)


#View mean COVID cases and mean COVID test conducted by settlement types


collapsed_dat_settlement_types <- combined_dat_ny_covid_pop %>%
  group_by(settlement_type) %>%
  summarise(Mean_COVID_cases = mean(sum_cumulative_no_of_positive),
            Mean_Tests_Conducted = mean(sum_cumulative_no_of_tests),
            Mean_Absolute_Population = mean(pop2022),
            Mean_Settlement_Density = mean (density_permilessquared))

collapsed_dat_settlement_types #View table

```

### Graphical Exploration

```{r, message=FALSE, warning=FALSE, fig.cap= "Cumulative Number of Positive COVID-19 Tests", fig.align='left', fig.height= 2.5 }

# Visualizing Distribution of Variables

###################### COVID-19 Cases ##############################

ggplot(data = combined_dat_ny_covid_pop, aes(sum_cumulative_no_of_positive))+
  geom_histogram() +
  xlab ("Cumulative Number of Positive COVID-19 Tests") +
  ggtitle("Histogram of Cumulative Number of Positive COVID-19 Tests")

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Log of Cumulative Number of Positive COVID-19 Tests", fig.align='left', fig.height= 2.5 }

## Log Transform

ggplot(data = combined_dat_ny_covid_pop, aes(log(sum_cumulative_no_of_positive))) +
  geom_histogram() +
  xlab ("Log of Cumulative Number of Positive COVID-19 Tests") +
  ggtitle("Histogram of Log of Cumulative Number of Positive COVID-19 Tests")

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Cumulative Number of COVID-19 Tests Conducted", fig.align='left', fig.height= 2.5 }

####################### Number of COVID Tests Conducted #######################

ggplot(data = combined_dat_ny_covid_pop, aes(sum_cumulative_no_of_tests))+
  geom_histogram() +
  xlab ("Cumulative Number of COVID-19 Tests Conducted") +
  ggtitle("Histogram of Cumulative Number of COVID-19 Tests Conducted")

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Log of Cumulative Number of Positive COVID-19 Tests Conducted", fig.align='left', fig.height= 2.5 }

## Log Transform

ggplot(data = combined_dat_ny_covid_pop, aes(log(sum_cumulative_no_of_tests))) +
  geom_histogram() +
  xlab ("Log of Cumulative Number of COVID-19 Tests Conducted") +
  ggtitle("Histogram of Log of Cumulative Number of COVID-19 Tests Conducted")

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Population of Counties Across New York State", fig.align='left', fig.height= 2.5 }

######################### County Population ##########################

ggplot(data = combined_dat_ny_covid_pop, aes(pop2022)) +
  geom_histogram() +
  xlab("Absolute Population Size of Counties Across New York State 2022") +
  ggtitle ("Histogram of Population of Counties Across New York State")

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Log of Population of Counties Across New York State", fig.align='left', fig.height= 2.5 }

## Log Transform

ggplot(data = combined_dat_ny_covid_pop, aes(log(pop2022))) +
  geom_histogram() +
  xlab("Log of Absolute Population of Counties Across New York State 2022 ") +
  ggtitle ("Log of Histogram of Population of Counties Across New York State")

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Density of Counties Across New York State", fig.align='left', fig.height= 2.5 }

################################## County Density ############################

ggplot(data = combined_dat_ny_covid_pop, aes(density_permilessquared)) +
  geom_histogram() +
  xlab("Density of Counties Across New York State") +
  ggtitle (" Histogram of Density of Counties Across New York State")

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Log of Density of Counties Across New York State", fig.align='left', fig.height= 2.5 }

## Log Transform

ggplot(data = combined_dat_ny_covid_pop, aes(log(density_permilessquared))) +
  geom_histogram() +
  xlab("Log of Density of Counties Across New York State") +
  ggtitle (" Log of Histogram of Density of Counties Across New York State")

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Log of Absolute Population Size of Counties by Settlement Type", fig.align='left', fig.height= 2.5 }
########### Visualize Distribution of Predictor Variables by Settlement Type

ggplot(data = combined_dat_ny_covid_pop, aes(x= settlement_type, y= log(density_permilessquared), fill = settlement_type))+
  geom_boxplot() +
  ggtitle ("Log of Absolute Population Size of Counties by Settlement Type") +
  xlab("Settlement Types") +
  ylab("Log of Absolute Population Size 2022")

```

```{r, message=FALSE, warning=FALSE, fig.cap= "Log of Population Density of Counties by Settlement Type", fig.align='left', fig.height= 2.5 }

ggplot(data = combined_dat_ny_covid_pop, aes(x= settlement_type, y= log(pop2022), fill = settlement_type))+
  geom_boxplot() +
    ggtitle ("Log of Population Density of Counties by Settlement Type") +
  xlab("Settlement Types") +
  ylab("Log of Absolute Population Size 2022")

```


### Spatial Context of Data

###```{r, message=FALSE, warning=FALSE}


# Read in counties shapefile into an sf dataframe, filtering for just NY counties (GURUBANDAA)

uscounty <- st_read('./Data/Spatial_Files/cb_2018_us_county_20m.shp', stringsAsFactors = TRUE) %>%
  filter(
    STATEFP == 36
    ) %>%
  st_as_sf(coords = c('SITE_LONGITUDE','SITE_LATITUDE'), crs=4269)


# Visualize data
counties.sf <- ggplot(data = uscounty) +
  geom_sf()

# Plot the data.
ggplot(data = counties.sf) +
  geom_sf()

# The EPSG code of the New York counties dataset is 4269. This is a geographic coordinate reference system, so it uses angular coordinate units. This CRS (4269) is associated with the datum NAD83.

# Reveal the names of the columns
colnames(combined_dat_ny_covid_pop)

# Join the flow data to our NWIS gage location spatial dataframe.
combined_dat_join <- merge(x = uscounty,
                          y = combined_dat_ny_covid_pop,
                          by.x = "NAME",
                          by.y = "County")

# Show the column names of the joined dataset.
colnames(combined_dat_join)

# Show the dimensions of this joined dataset.
dim(combined_dat_join)

# Map the population of each New York county in 2022.
mapview(combined_dat_join, 
        zcol = 'pop2022', 
        col.regions = brewer.pal(2, 'RdBu'))

# Map the settlement types of each New York county in 2022.
mapview(combined_dat_join, 
        zcol = 'settlement_type', 
        col.regions = brewer.pal(2, 'RdBu'))

# Map the sum of the cumulative number of tests administered in each New York county in 2022.
mapview(combined_dat_join, 
        zcol = 'sum_cumulative_no_of_tests', 
        col.regions = brewer.pal(2, 'RdBu'))

# Map the sum of the cumulative number of positive test results in each New York county in 2022.
mapview(combined_dat_join, 
        zcol = 'sum_cumulative_no_of_positive', 
        col.regions = brewer.pal(2, 'RdBu'))


```

### Export Processed data

```{r, message=FALSE, warning=FALSE}

getwd() #Check working directory

#Full processed data set

write.csv2(combined_dat_ny_covid_pop,"./Data/Processed_Data/Combined_NY_COVID_Pop.csv")

#Collapsed data set

write.csv2(collapsed_dat_settlement_types,"./Data/Processed_Data/collapsed_dat_settlement_types.csv")
```


\newpage


# Analysis

As a central question of this study is "How  well does population density and
absolute population predict COVID-19 cases?", the predictor variables are 
population density per miles squared and absolute population while the outcome 
variable is sum cumulative positive number of positive COVID-19 tests which is 
being used as a proxy for COVID-19 cases. Our unit of analysis is counties and 
settlement types depending on the primary objective of 
each sub-question. Settlement types are classified based on our definitions of
rural, micropolitan, and urban settlement types. Our hypotheses for each study question
is indicated below.

## Question 1:  What is the correlation between population density and absolute population and the number of COVID-19 cases(cumulative number of positive tests) across New York State?

### Absolute Population and COVID-19 Cases

#### Hypotheses:

Null Hypothesis: As the ranks of absolute population change, the ranks of the number of COVID-19 cases do not change. 
Alternative Hypothesis:  As the ranks of absolute population change, the ranks of the number of COVID-19 cases change.


#### Assumptions:

1. The relationship between both variables is monotonic.


```{r, message=FALSE, warning=FALSE}

# Spearman rank correlation test as distributions are skewed


correlation_results_absol_pop <- cor.test(x= log(combined_dat_ny_covid_pop$pop2022),
                                          y = log(combined_dat_ny_covid_pop$sum_cumulative_no_of_positive), method = "spearman")
correlation_results_absol_pop

# Visualize plot (Log-transformed for better visualization)

ggplot(combined_dat_ny_covid_pop, aes(x = log(pop2022), y = log(sum_cumulative_no_of_positive)))+
  geom_point() +
  ylab ("Log of COVID-19 Cases") +
  xlab ("Log of Absolute Population Across Counties") +
  ggtitle ("The Correlation Between Absolute Population 
           Across Counties and COVID-19 Cases") +
  stat_smooth (method = 'lm') +
  annotate("text",x=9,y=20, label=(paste0("R=0.99")))
  
```

### Population Density and COVID-19 Cases

#### Hypotheses:

Null Hypothesis: As the ranks of population density change, the ranks of the number of COVID-19 cases do not change.
Alternative Hypothesis: As the ranks of population density change, the ranks of the number of COVID-19 cases change.

#### Assumptions:

1. The relationship between both variables is monotonic.

```{r, message=FALSE, warning=FALSE}

# Spearman rank correlation test as distributions are skewed


correlation_results_pop_density <- cor.test(x= log(combined_dat_ny_covid_pop$density_permilessquared),
                                          y = log(combined_dat_ny_covid_pop$sum_cumulative_no_of_positive), method = "spearman")
correlation_results_pop_density

# Visualize plot (Log-transformed for better visualization)

ggplot(combined_dat_ny_covid_pop, aes(x = log(density_permilessquared), y = log(sum_cumulative_no_of_positive)))+
  geom_point() +
  xlab (" Log of Population Density (per miles squared)") +
  ylab ("Log of COVID-19 Cases") +
  ggtitle ("The Correlation Between Population Densities
           Across Counties and COVID-19 Cases") +
  geom_smooth (method = 'lm')+
  annotate("text",x=3,y=20, label=(paste0("R=0.89")))
```


## Question 2:  What is the relationship between the predictor variable: settlement types and the response variable: cumulative COVID-19 cases?

### Negaative Binomial Regression

A negative binomial regression is used as the outcome variable can be classified as
over-dispersed count data. 

#### Hypotheses:

Null Hypothesis: There is no trend between the response and outcome variable.
Alternative Hypothesis: There is a trend between the response and outcome variables.

#### Assumptions:

1. The response variable (COVID cases) is a count variable.
2. Each observation of data is independent of each other.
3. The mean and variance are not equal.


```{r, message=FALSE, warning=FALSE}

## Negative Binomial Regression

summary (modelnb <- glm.nb(sum_cumulative_no_of_positive ~ factor(settlement_type), data = combined_dat_ny_covid_pop))


par (mfrow = c(2,2))
plot(modelnb) #Check fit


## Check our assumption that a negative binomial is a better model than the poisson model

summary (modelpoisson <- glm(sum_cumulative_no_of_positive ~ factor(settlement_type), family = "poisson", data = combined_dat_ny_covid_pop))

##Residual deviance is way larger than degrees of freedom (DF). Negative binomial plot better.


```


## Question 3: What is the correlation between absolute population and population density and the number of COVID-19 tests done?

### Absolute Population and COVID-19 Cases

#### Hypotheses:

Null Hypothesis: As the ranks of absolute population change, the ranks of the number of COVID-19 tests conducted do not change. 
Alternative Hypothesis:  As the ranks of absolute population change, the ranks of the number of COVID-19 tests conducted change.

#### Assumptions:

1. The relationship between both variables is monotonic.

```{r, message=FALSE, warning=FALSE}

# Spearman rank correlation test as distributions are skewed


correlation_results_absol_pop <- cor.test(x= log(combined_dat_ny_covid_pop$pop2022),
                                          y = log(combined_dat_ny_covid_pop$sum_cumulative_no_of_tests), method = "spearman")
correlation_results_absol_pop

# Visualize plot (Log-transformed for better visualization)

ggplot(combined_dat_ny_covid_pop, aes(x = log(pop2022), y = log(sum_cumulative_no_of_tests)))+
  geom_point() +
  ylab ("Log of COVID-19 Tests Conducted") +
  xlab ("Log of Absolute Population Across Counties") +
  ggtitle ("The Correlation Between Absolute Population 
           Across Counties and COVID-19 Tests Conducted") +
  stat_smooth (method = 'lm') +
  annotate("text",x=9,y=22, label=(paste0("R=0.97")))
  
```

### Population Density and COVID-19 Cases

#### Hypotheses:

Null Hypothesis: As the ranks of population density change, the ranks of the number of COVID-19 tests conducted do not change.
Alternative Hypothesis: As the ranks of population density change, the ranks of the number of COVID-19 tests conducted change.

#### Assumptions:

1. The relationship between both variables is monotonic.

```{r, message=FALSE, warning=FALSE}

# Spearman rank correlation test as distributions are skewed


correlation_results_pop_density <- cor.test(x= log(combined_dat_ny_covid_pop$density_permilessquared),
                                          y = log(combined_dat_ny_covid_pop$sum_cumulative_no_of_tests), method = "spearman")
correlation_results_pop_density

# Visualize plot (Log-transformed for better visualization)

ggplot(combined_dat_ny_covid_pop, aes(x = log(density_permilessquared), y = log(sum_cumulative_no_of_tests)))+
  geom_point() +
  xlab (" Log of Population Density (per miles squared)") +
  ylab ("Log of COVID-19 Tests Conducted") +
  ggtitle ("The Correlation Between Population Densities
           Across Counties and COVID-19 Tests Conducted") +
  geom_smooth (method = 'lm')+
  annotate("text",x=3,y=23, label=(paste0("R=0.86")))

```


###  Question 3b: Does the stronger correlated relationship vary among settlement types?

### Absolute Population and Settlement Types

```{r, message=FALSE, warning=FALSE}

### Correlation Plots by Settlement Type

##Urban

Urban_dat <- combined_dat_ny_covid_pop %>% 
  filter(settlement_type=="Urban")


correlation_results_absolute_pop_urban <- cor.test(x= log(Urban_dat$pop2022),
                                          y = log(Urban_dat$sum_cumulative_no_of_tests), method = "spearman")

correlation_results_absolute_pop_urban

## Micropolitan

Micropolitan_dat <- combined_dat_ny_covid_pop %>% 
  filter(settlement_type=="Micropolitan")


correlation_results_absolute_pop_microp <- cor.test(x= log(Micropolitan_dat$pop2022),
                                          y = log(Micropolitan_dat$sum_cumulative_no_of_tests), method = "spearman")

correlation_results_absolute_pop_microp

##Rural

Rural_dat <- combined_dat_ny_covid_pop %>% 
  filter(settlement_type=="Rural")


correlation_results_absolute_pop_rural <- cor.test(x= log(Rural_dat$pop2022),
                                          y = log(Rural_dat$sum_cumulative_no_of_tests), method = "spearman")

correlation_results_absolute_pop_rural
```

### Visualization of 3b

```{r, message=FALSE, warning=FALSE}

ggplot(combined_dat_ny_covid_pop, aes(x = log(pop2022), y = log(sum_cumulative_no_of_positive)))+
  geom_point() + 
  ylab ("Log of Cumulative Number of COVID_19 Cases") +
  xlab ("Absolute Population (2022)")+
  facet_wrap(vars(settlement_type)) 
 
```


\newpage

# Summary and Conclusions


\newpage

# References

1. CDC. (2020, February 11). COVID-19 and Your Health. Centers for Disease Control and Prevention. https://www.cdc.gov/coronavirus/2019-ncov/your-health/reinfection.html

2. CDCMMWR. (2020). Geographic Differences in COVID-19 Cases, Deaths, and Incidence—United States, February 12–April 7, 2020. MMWR. Morbidity and Mortality Weekly Report, 69. https://doi.org/10.15585/mmwr.mm6915e4

3. Coronavirus. (n.d.). Retrieved March 14, 2022, from https://www.who.int/westernpacific/health-topics/coronavirus

4. Hamidi, S., Sabouri, S., & Ewing, R. (2020). Does Density Aggravate the COVID-19 Pandemic? Journal of the American Planning Association, 86(4), 495–509. https://doi.org/10.1080/01944363.2020.1777891

5. New York Population 2022 (Demographics, Maps, Graphs). (n.d.). Retrieved December 10, 2022, from https://worldpopulationreview.com/states/new-york-population

6. New York State Tracking Program | Tracking | NCEH | CDC. (2019, February 20). https://www.cdc.gov/nceh/tracking/profiles/New_York_State_Profile.htm

7. Settlements Overview & Types | What are Settlements? - Video & Lesson Transcript. (n.d.). Study.Com. Retrieved December 10, 2022, from https://study.com/learn/lesson/settlements-overview-types.html

8. Times, T. N. Y. (2020, April 1). New York Coronavirus Map and Case Count. The New York Times. https://www.nytimes.com/interactive/2021/us/new-york-covid-cases.html

9. Truelove, S., Smith, C. P., Qin, M., Mullany, L. C., Borchering, R. K., Lessler, J., Shea, K., Howerton, E., Contamin, L., Levander, J., Kerr, J., Hochheiser, H., Kinsey, M., Tallaksen, K., Wilson, S., Shin, L., Rainwater-Lovett, K., Lemairtre, J. 

10. C., Dent, J., … Viboud, C. (n.d.). Projected resurgence of COVID-19 in the United States in July—December 2021 resulting from the increased transmissibility of the Delta variant and faltering vaccination. ELife, 11, e73584. https://doi.org/10.7554/eLife.73584

11. USDA ERS - What is Rural? (n.d.). Retrieved November 27, 2022, from https://www.ers.usda.gov/topics/rural-economy-population/rural-classifications/what-is-rural.aspx

12. What Is Coronavirus? (2022, July 29). https://www.hopkinsmedicine.org/health/conditions-and-diseases/coronavirus

13. Wong, D. W. S., & Li, Y. (2020). Spreading of COVID-19: Density matters. PLOS ONE, 15(12), e0242398. https://doi.org/10.1371/journal.pone.0242398


