---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: "Tracking COVID-19 Cases Across Different Geographic Neighborhoods in New York State"
subtitle: "Web address for GitHub repository"
author: "SMiltonDZouGKhalsa"
date: "2022-11-22"
fontsize: 12pt
mainfont: Times New Roman
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

## Set-up Data Set

```{r, message=FALSE}

#Set working directory


#Load packages

library(rvest)
library(tidyverse)
library(ggplot2)
library (lubridate)
library (dplyr)


#Set theme

my_theme <- theme_bw(base_size = 12) + 
  theme(axis.text = element_text(color = "black"), 
      legend.position = "top", legend.justification = "center") +
  theme(plot.title = element_text(hjust = 0.5))
theme_set(my_theme)

```


## Import and Merge Data Sets

Important things to note: (delete later)

-There are 62 counties in New York State.

```{r, echo = FALSE}

#Import NY State county level COVID data (March 1, 2020 to November 22, 2022) from NY Health.

ny_covid_dat <- read.csv("./Data/Raw_Data/NY_COVID_County_Level.csv", stringsAsFactors = TRUE)

ny_covid_dat #View Data

#Import NY population data (2022)

ny_pop_dat <- read.csv("./Data/Raw_Data/NY_County_Level_Population_Data.csv", stringsAsFactors = TRUE)

ny_pop_dat

#Taken from https://www.census.gov/library/stories/state-by-state/new-york-population-change-between-census-decade.html

#Merge data sets

combined_dat_ny_covid_pop <- ny_pop_dat %>% 
  left_join(ny_covid_dat, by = "County")

### Explore Data

head(combined_dat_ny_covid_pop)
summary(combined_dat_ny_covid_pop)#View Data

#Change data types
colnames(combined_dat_ny_covid_pop)
combined_dat_ny_covid_pop$sum_cumulative_no_of_positive <- as.numeric(combined_dat_ny_covid_pop$sum_cumulative_no_of_positive)
combined_dat_ny_covid_pop$sum_new_positives <- as.numeric(combined_dat_ny_covid_pop$sum_new_positives)
combined_dat_ny_covid_pop$sum_total_number_of_tests <- as.numeric(combined_dat_ny_covid_pop$sum_total_number_of_tests)
combined_dat_ny_covid_pop$sum_cumulative_no_of_tests <- as.numeric(combined_dat_ny_covid_pop$sum_cumulative_no_of_tests)
combined_dat_ny_covid_pop$median_test_positive <- as.numeric(combined_dat_ny_covid_pop$median_test_positive)

summary(combined_dat_ny_covid_pop)  #View Data

#Export Processed data

getwd() #Check working directory

write.csv2(combined_dat_ny_covid_pop,"./Data/Processed_Data/Combined_NY_COVID_Pop.csv")
```

# Rationale and Research Questions



\newpage

# Dataset Information



\newpage

# Exploratory Analysis 



\newpage

# Analysis



## Question 1: <insert specific question here and add additional subsections for additional questions below, if needed>

## Question 2: 




\newpage

# Summary and Conclusions


\newpage

# References
<add references here if relevant, otherwise delete this section> 
